Evaluating sensitivity to measurement error in dissolved gas measurements

### 2024-04-05 scripts/DG_sensitivity_to_measurement_error.Rmd
adding uncertainty in denominator

```
# Define GC measurement error
Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.

@@

# Simulate measurement error for gas measured in water
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector

@@

# Simulate measurement error for gas measured in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

@@

# Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

@@

# Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

@@

# Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

@@

# Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

@@

# Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

@@

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me* 1e9) * 2 # +/- 0.302 nmol N2O
```

# Incorporate measurement uncertainty into source/sink status
The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.  The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 2`nM.  Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 2`nM of the equilibrium value cannot be determined as a source or a sink.

@@

dg <- dg %>%
  mutate(n2o.src.snk.with.error = case_when(
    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 2 ~ n2o.src.snk,
    TRUE ~ "undetermined"))

@@

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.1, 10), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

>>>>

```
## Define GC measurement error
Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.

@@

## Simulate measurement error in GC measurement of N2O in equilibrated headspace
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector

@@

## Simulate measurement error for GC measurement of N2O in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

@@

## Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

@@

## Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

@@

## Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

@@

## Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

@@

## Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

@@

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me* 1e9) * 1.96 # +/- 0.302 nmol N2O
```

## Incorporate measurement uncertainty into source/sink status
The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.  The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 1.96`nM.  Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 1.96`nM of the equilibrium value cannot be determined as a source or a sink.

@@

dg <- dg %>%
  mutate(n2o.src.snk.with.error = case_when(
    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 1.96 ~ n2o.src.snk,
    TRUE ~ "undetermined"))

@@

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.1, 10), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```
# Set up model for simulating true saturation concentration

The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water that is in equilibrium with the atmosphere is:

$$C = 1^{-6}B*C_{air}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15})$$
```{r c_sat}
C_sat <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))

C_sat <- C_sat * 1e9 #convert mol to nmol

## Simulate measurement error in GC measurement of N2O in air
```{r simulate_me_sat_C_a, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_me[i] * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_me = C_a_me)

ggplot(C_sat_df, aes(x = C_a_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (ppm)")

ggplot(C_sat_df, aes(x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes(x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in water temperature
```{r simulate_me_sat_T_c, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, T_c_me = T_c_me)

ggplot(C_sat_df, aes(x = T_c_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = T_c , color = 'red' ) + 
  ggtitle( "Simulated measurement error in temperature") +
  xlab( "temperature (C)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline(xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error" ) +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in barometric pressure
```{r simulate_me_sat_B, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, B_me = B_me)

ggplot(C_sat_df, aes( x = B_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error for all observables
```{r simulate_me_sat_all, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_me[ i ] * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat)

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "absolute error (nMol)")

# standard deviation of a measurement is:
sd(C_me_sat* 1e9) # 0.20 nmol L-1

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me_sat * 1e9) * 1.96 # +/- 0.395 nmol N2O
```
